---
output: github_document
always_allow_html: true
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  dev = "svg" # set output device to svg
)
```

# spectralem
[![R-CMD-check](https://github.com/jan-meissner/spectralem/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/jan-meissner/spectralem/actions/workflows/R-CMD-check.yaml)
[![codecov](https://codecov.io/github/jan-meissner/spectralem/branch/main/graphs/badge.svg)](https://app.codecov.io/gh/jan-meissner/spectralem/commits?branch=main)

## Installation

1. If `RcppFaddeeva` is not available for the current R version use:

```r
remotes::install_github("cran/RcppFaddeeva")
```

2. You can install the development version with:

```r
remotes::install_github("jan-meissner/spectralem")
```

## Examples

Fitting 30 Peaks:

```{r include = FALSE}
library(ggplot2)
```
```{r}
library(spectralem)

data <- synthetic.signal(seed = 1, K = 30, noise = 0.001)
x <- data$x
y <- data$y

res <- spectralem(x, y, K = 30)

# get position of fitted peaks
res$fit_params$pos

# plot the fit
library(ggplot2)
pd <- data.frame(x = x, y = y, fit = voigt.model(x, res$fit_params))
ggplot(pd, aes(x)) +
  geom_line(aes(y = y, colour = "y")) +
  geom_line(aes(y = fit, colour = "fit")) +
  theme_bw()
```

## Funding

This work was developed under a Seed Fund Project (2021) of the RWTH
Aachen University, funded under "the Excellence Strategy of the Federal
Government and the Länder".
Project: "Spectra-Bayes: A Bayesian statistical machine learning model
for spectral reconstruction" (Project Leaders: Prof. Dr. Maria Kateri,
Prof. Dr.-Ing. Hans-Jürgen Koß)


