---
title: "Get started"
author: "Jan Meißner"
output:
  rmarkdown::html_vignette:
    fig_caption: yes
vignette: >
  %\VignetteIndexEntry{Get started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
fig_width: 6
fig_height: 12
---
<!-- Add fullscreen to plotly -->
<script type="text/javascript">
  function jump_header(key){
    $(':header:contains('+key+')')[0].scrollIntoView();
  }

  function jump_marked(key){
    $('#' + key)[0].scrollIntoView();
  }

  function plotZoom(el){
      if(document.fullscreen) {
        document.exitFullscreen()
      } else {
        $(el).closest('.js-plotly-plot')[0].requestFullscreen();
      }
  }

  $( document ).ready(function() {
    $(".modebar-btn.plotlyjsicon.modebar-btn--logo").replaceWith(
    `
    <a rel="tooltip" onclick=plotZoom(this) class="modebar-btn fullscreen-btn" data-title="Full Screen" data-attr="zoom" data-val="auto" data-toggle="false" data-gravity="n" >
      <svg xmlns="http://www.w3.org/2000/svg" viewBox="0 0 448 512" class="icon" height="1em" width="1em">
        <!--! Font Awesome Pro 6.1.1 by @fontawesome - https://fontawesome.com License - https://fontawesome.com/license (Commercial License) Copyright 2022 Fonticons, Inc. -->
        <path d="M128 32H32C14.31 32 0 46.31 0 64v96c0 17.69 14.31 32 32 32s32-14.31 32-32V96h64c17.69 0 32-14.31 32-32S145.7 32 128 32zM416 32h-96c-17.69 0-32 14.31-32 32s14.31 32 32 32h64v64c0 17.69 14.31 32 32 32s32-14.31 32-32V64C448 46.31 433.7 32 416 32zM128 416H64v-64c0-17.69-14.31-32-32-32s-32 14.31-32 32v96c0 17.69 14.31 32 32 32h96c17.69 0 32-14.31 32-32S145.7 416 128 416zM416 320c-17.69 0-32 14.31-32 32v64h-64c-17.69 0-32 14.31-32 32s14.31 32 32 32h96c17.69 0 32-14.31 32-32v-96C448 334.3 433.7 320 416 320z"/>
      </svg>
    </a>
    `);
  });
</script>
```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE
)
set.seed(1234)
```
The R package **spectralem** uses a EM algorithm <paperref> to fit a model of Voigt profiles to a signal:
$$
f(x) = a x + b + \sum_{j=1}^K \beta_j V(x \mid \theta_j) + \sum_{i=1} \alpha_i f_i(x)
$$
Here $\beta_j > 0$ are the amplitudes of the Voigt profiles $V$ where $\theta_j$ represents the position,
Gaussian and Lorentzian width of the voigt profile. Additionally, it is possible to fit arbitrary positive background functions \eqn{f_i(x) > 0}, these are by default zero.
Their amplitudes are constrained to $\alpha_i > 0$.

The algorithm is fast and even for 30 peaks takes only roughly a minute to fit.

## General Usage

```{r, fig.width=8, fig.height=4}
library(spectralem)

data <- synthetic.signal(seed = 1, K = 30, noise = 0.005)
x <- data$x
y <- data$y

res <- spectralem(x, y, K = 30, print_progress = FALSE) # no progress bar for rmd

# get position of fitted peaks
res$fit_params$pos

fit <- voigt.model(x, res$fit_params)
# or alternatively fit <- res$fit


# plot the fit
library(plotly)
fig <- plotly::plot_ly(x = x, y = y, name = 'true', type = 'scatter', mode = 'lines')
fig <- fig %>% plotly::add_lines(y = fit, name = 'fit', type = 'scatter', mode = 'lines')
fig
```
## Passing optional start parameters

Can be done by passing the additional list 'start_peaks'.
```{r fig.width=8, fig.height=4}
data <- synthetic.signal(1212, 4, 0.001)

res <- spectralem(
  data$x, data$y,
  K = 4,
  start_peaks = list(
    pos = c(465, 644, 828, 862),
    gwidth = c(5, 5, 5, 5),
    lwidth = c(0.1, 0.1, 0.1, 0.1)
  ),
  print_progress = FALSE # no progress bar for rmd
)

# plot the fit
library(plotly)
fig <- plotly::plot_ly(x = data$x, y = data$y, name = 'true', type = 'scatter', mode = 'lines')
fig <- fig %>% plotly::add_lines(y = res$fit, name = 'fit', type = 'scatter', mode = 'lines')
fig
```
## Using BSplines as a background model

Using 'background_model' any fixed background signal can be added. It is not advised to set 'linear = FALSE'.
```{r fig.width=8, fig.height=4}
data <- synthetic.signal(seed = 312, K = 4, noise = 0.005)
x <- data$x
y <- data$y

# add spline background to data
library(splines2)
bsMat <- bSpline(x, df=3)
f1 <- c(bsMat[, 1])
f2 <- c(bsMat[, 2])
f3 <- c(bsMat[, 3])
y <- y + 0.4*f1 + 0.04*f2 + 0.2*f3

# fit
res <- spectralem(x, y, K = 4, background_model = list(linear=TRUE, f1, f2, f3), print_progress = FALSE)

# voigt.model does not support non linear background; add them manually
alpha <- res$fit_params$background_amps
fit <- voigt.model(x, res$fit_params) + f1 * alpha[1] + f2 * alpha[2] + f3 * alpha[3]
# or alternatively fit <- res$fit

# plot the fit
library(plotly)
fig <- plotly::plot_ly(x = x, y = y, name = 'true', type = 'scatter', mode = 'lines')
fig <- fig %>% plotly::add_lines(y = fit, name = 'fit', type = 'scatter', mode = 'lines')
fig
```
