---
title: "Get started"
author: "Jan Meißner"
output:
  rmarkdown::html_vignette:
    fig_caption: yes
vignette: >
  %\VignetteIndexEntry{Get started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
fig_width: 6
fig_height: 12
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE,
  fig.path = "man/figures/GETSTARTED-",
  out.width = "100%",
  dev = "svg" # set output device to svg
)
set.seed(1234)
```
The R package **spectralem** uses a EM algorithm <paperref> to fit a model of Voigt profiles to a signal:
$$
f(x) = a x + b + \sum_{j=1}^K \beta_j V(x \mid \theta_j) + \sum_{i=1} \alpha_i f_i(x)
$$
Here $\beta_j > 0$ are the amplitudes of the Voigt profiles $V$ where $\theta_j$ represents the position,
Gaussian and Lorentzian width of it. Additionally, it is possible to fit arbitrary positive background functions \eqn{f_i(x) > 0}, these are by default zero.
Their amplitudes are constrained to $\alpha_i > 0$.

The algorithm takes roughly a minute to fit 30 peaks.

## General Usage
```{r, fig.width=8, fig.height=4}
library(spectralem)

data <- synthetic.signal(seed = 1, K = 30, noise = 0.005)
x <- data$x
y <- data$y

res <- spectralem(x, y, K = 3, print_progress = FALSE) # no progress bar for vignettes

# get position of fitted peaks
res$fit_params$pos

fit <- voigt.model(x, res$fit_params)
# or alternatively fit <- res$fit

# plot the fit
library(ggplot2)
pd <- data.frame(x = x, y = y, fit = fit)
ggplot(pd, aes(x)) +
  geom_line(aes(y = y, colour = "y")) +
  geom_line(aes(y = fit, colour = "fit")) +
  theme_bw()
```